abxD01 Paper

### 2015-03-11 PAPER.Rmd
Add to discussion

Through our antibiotic perturbation experiments we identified a panel of bacteria with either a strong positive or negative relationship with subsequent levels of *C. difficile* colonization. A large number of *Porphyromonadaceae*, *Lachnospiraceae*, *Ruminococcaceae*, other *Clostridia*, and *Alistipes* were found to have strong potential resistance abilities against *C. difficile*. Conversely, several members of the *Proteobacteria* phylum, including *Pseudomonas*, *Streptococcus*, and especially *Escherichia* have strong potential positive influences on *C. difficile* colonization. These results were supported by their significant correlation values across several antibiotic regimens. By either reducing the amount of microbiota perturbation or allowing recovery following perturbation, we were able to capture microbiota structures which had relatively increased resistance to *C. difficile* colonization. Using this collective microbiota data spanning a wide range of colonization resistance levels, we next moved beyond correlation to prediction of *C. difficile* colonization based on the composition of the microbiota. Through modeling we identified bacteria within the community which contributed greatest to accurately predicting *C. difficile* levels. This is the first analysis both measuring and predicting colonization resistance levels based on the microbiota structure.

Correlation analysis identified whole families with significant negative or positive relationships with *C. difficile*.  Many of the correlaions we observed were not surprising given other studies looking at the relationship between *C. difficile* and the microbiome. For instance, monocolonization of germfree mice with either a *Lachnospiraceae* species or *E. coli* were tested for their resistance to *C. difficile* colonization [Reeves]. Lachnospiraceae was able to decrease the levels of *C. difficile* observed, while *E. coli* had similar colonization levels compared to germfree mice given *C. difficile* alone. Association studies in humans with and without CDI in hospitals have also highlighted several bacteria found in this study, including *Lachnospiraceae*, *Rumicoccaceae*, *Clostridia*, and *Alistipes* as decreased and *E. coli* and *Streptococcus* as increased in subjects with CDI [Schubert, Antharam]. While we observe overlaps in bacteria with potentially protective roles in both mice and humans at broad taxonomy levels, there are notable differences at refined levels. *Porphyromonadaceae*, for example, are more common among mice than humans, where instead there are strong negative associations of *C. difficile* with *Bacteroides* [Schubert]. However, correlation analysis alone was insuficient in identifying all informative bacteria. Random forest modeling allowed us to identify several bacterial OTUs that did not have significant correlations with *C. difficile* yet improved model accuracy. *Turicibacter* and *Akkermansia*, for example, both had non-significant correlations with *C .difficile* yet were more important for model accuracy over OTUs with stronger correlations (*Figure 5BC*). While these bacteria may not have important direct effects on *C. difficile*, they may interact positively or negatively with bacteria that do have strong relationships with *C. difficile*. Few studies attempt to model the microbiome to predict disease or estimate health [Pamer, clostridium paper, ].

@@

o  How relates to other studies
•	Antonopoulos et al-reproducible community dynamics
•	Antunes et al (finlay)—abx treatment on intestinal metabalome
•	Sekirov—antibiotic induced perturbations of intestinal microbiota
•	 “Change in the compostion alone of the intestinal microbiota (in the absence of significant changes in total numbers of intestinal microbes) prior to Salmonella serovar Typhimurium infection was sufficient to make mice more susceptible to Salmonella serovar Typhimurium, indicating that different subsets of the microbiota may play protective roles or enhance susceptibility to enteric infections.”
•	Streptomycin and vancomycin doses were administered to mice prior to challenge with another gastrointestinal pathogen Salmonella enterica serovar Typhymurium (Sekirov 2008).  This study showed ___. Changes in community composition were dose dependent.
•	Strep increases bacteroidetes (CFB), vanc increases firms and then proteo (with increasing dose)
•	Reeves/theriot/robinson—clindamycin/cef/cocktail/antbiotic administration alters the community structure
•	Croswell—prolonge impact of abx on intestinal microbial ecology salmonella
•	Rea- effect of  broad and narrow spectrum antimicrobials
•	Ferreira—intestinal microbiota plays a role in salmonella induced colitis
•	Reeves with clone libraries and cef
•	Theriot?—at least the metabolome paper—before and after with metabolites and cef, also does she look at anything taxonomically? yup

Many of the relationships we observed in our study between specific bacteria and *C. difficile* are supported in the literature. –see my comments in this section
Within the Porphyromonadaceae, only OTU505 had a significant positive relationship with subsequent *C. difficile* colonization (ρ= +0.21), which by blastn shares 99% identity with Coprobacter fastidiosus. Similarly within the Lachnospiraceae family, only one OTU had a positive relationship with *C. difficile* colonization (OTU174, ρ= +0.22). The Lachnospiraceae group also contained an OTU that had an exceptionally strong negative correlation (OTU39, ρ= +0.77). The Clostridia class was largely composed of the order Clostridiales (n=6), of which only one had a positive relationship with *C. difficile* colonization (OTU154, ρ= +0.26).

Several broader groups of bacteria can work together (in some combination) to provide CR …possible that proteobacteria in general are bad …These results indicate that many related species within the larger/broader bacterial groups may be able to serve in providing colonization resistance against *C. difficile*.
Other taxons in Figure 3 had more species specific correlations. For instance the Actinobacteria phylum included Enterorhabdus (OTU29, ρ= -0.70), Coriobacteriaceae (OTU50, ρ= -0.50), Bifodbacterium (OTU22, ρ= -0.34), Actinomyces (OTU58, ρ= +0.23). The Lactobacillales group is divided into two negatively correlated Lactobacillus OTUs (OTUs 23, 21, ρ= -0.73, -0.40 respectively) and two positively correlated Streptococcus OTUs (OTUs 78, 79). Similarly, the Bacteroidales group included species at both extremes, for example, an Alistipes (OTU20, ρ= -0.78) and a Bacteroides species (OTU44, ρ= +0.34). [END sent, all the other bad bugs look more like random smaller level taxonomic classification divisions, as in it’s a genus/species level effect type deal]
--functional redundancy? Across related large groups like porphyros?
--mostly Firmicutes—better competitors with cdiff? Because similar nutrients resources?
Also functional redundancy with a highly diverse community… not just about who possibly but also ab being diverse—emphasizes community as a whole

@@

[using this time model for the metro… I wonder if it doesn’t work because of the time difference in the amount of recovery time? I’m picturing sort of an echo/ripple effect where at first the waves, or changes, are really large and then over time that rate of change of the community slows and the changes less pronounced/smaller. In this context the metro communities were at a stage farther along in the echo, meaning changes not happening as fast. Whereas the measurements we’ve taken and are using to make predictions are from the community at a time when you expect more/bigger/faster changes. So then the results are calculated assuming one rate but actually in a different (slower).
By this image, the initial strength of the perturbation matters too.]

--also somewhere, ecoli are typically on avg (__% of the community, typically wayyyy outnumbered… whats the avg relabund of all those CR OTUs in an untreated mouse, how does that compare to the abundance of those bacteria I’ve implicated in –CR) …point is that Proteobacteria are usually outnumbered… don’t have a chance, gut is suited to bact/firm (which not surprisingly make up the major players in CR).. I bet those other outlier bad bugs are also in low abund in untreated mice (?). but when the gut environment changes for w/e reason, can give those rarer bad ones/proteo/esp e. coli a chance to take hold. They could just be individually opportunistic or work synergistically with cdiff. If individually opportunistic, it might be that they are facing pretty much a common enemy which is the lach/rum/clos/porphs and end up having similar techniques/niches as a result of that common enemy. Working synergistically—eg. b. theta sialidase where it cleaves off resources for cdiff.

>>>>

Through our antibiotic perturbation experiments we identified a panel of bacteria with either strong positive or negative relationships with subsequent levels of *C. difficile* colonization. A large number of *Porphyromonadaceae*, *Lachnospiraceae*, *Ruminococcaceae*, other *Clostridia*, and *Alistipes* were found to have strong potential resistance abilities against *C. difficile*. Conversely, several members of the *Proteobacteria* phylum, including *Pseudomonas*, *Streptococcus*, and especially *Escherichia* have strong potential positive influences on *C. difficile* colonization. These results were supported by their significant correlation values across several antibiotic regimens. By either reducing the amount of microbiota perturbation or allowing recovery following perturbation, we were able to capture microbiota structures which had increased resistance to *C. difficile* colonization. Using this collective microbiota data spanning a wide range of colonization resistance levels, we next moved beyond correlation to prediction of *C. difficile* colonization based on the composition of the microbiota. Through modeling we identified bacteria within the community which contributed greatest to accurately predicting *C. difficile* levels. This is the first analysis both measuring and predicting colonization resistance levels based on the microbiota structure.

Correlation analysis identified major families with significant negative or positive relationships with *C. difficile*.  Many of the correlaions we observed were not surprising given other studies looking at the relationship between *C. difficile* and the microbiome. For instance, monocolonization of germfree mice with either a *Lachnospiraceae* species or *E. coli* were tested for their resistance to *C. difficile* colonization [Reeves]. Pre-colonization with *Lachnospiraceae* was able to decrease the levels of *C. difficile* observed, while *E. coli* had similar colonization levels compared to germfree mice given *C. difficile* alone. Association studies in humans with and without CDI in hospitals have also highlighted several bacteria found in this study, including *Lachnospiraceae*, *Rumicoccaceae*, *Clostridia*, and *Alistipes* as decreased and *E. coli* and *Streptococcus* as increased in subjects with CDI [Schubert, Antharam]. While we observe overlaps in bacteria with potentially protective roles in both mice and humans at broad taxonomy levels, there are notable differences at refined levels. *Porphyromonadaceae* (Class *Bacteroidales*), for example, are more common among mice than humans, where instead there are strong negative associations of *C. difficile* with *Bacteroides* (Class *Bacteroidales*) [Schubert].

Correlation analysis alone was insufficient in identifying all informative bacteria. Random forest modeling allowed us to identify several bacterial OTUs that did not have significant correlations with *C. difficile* yet improved model accuracy. *Turicibacter* and *Akkermansia*, for example, both had non-significant correlations with *C .difficile* yet were more important for model accuracy over OTUs with stronger correlations (*Figure 5BC*). While these bacteria may not have important direct effects on *C. difficile*, they may interact positively or negatively with bacteria that do have strong relationships with *C. difficile*. Modeling of the microbiome allows us to characterize *C. difficile* colonization in terms of a community, which is necessary to capture the complex relationships among its members. Few studies attempt to model the microbiome to predict disease or estimate health [Pamer, clostridium paper, psoriases, schubert].

@@

o  How relates to other studies
•	Antonopoulos et al-reproducible community dynamics
•	Antunes et al (finlay)—abx treatment on intestinal metabalome
•	Sekirov—antibiotic induced perturbations of intestinal microbiota
•	 “Change in the compostion alone of the intestinal microbiota (in the absence of significant changes in total numbers of intestinal microbes) prior to Salmonella serovar Typhimurium infection was sufficient to make mice more susceptible to Salmonella serovar Typhimurium, indicating that different subsets of the microbiota may play protective roles or enhance susceptibility to enteric infections.”
•	Streptomycin and vancomycin doses were administered to mice prior to challenge with another gastrointestinal pathogen Salmonella enterica serovar Typhymurium (Sekirov 2008).  This study showed ___. Changes in community composition were dose dependent.
•	Strep increases bacteroidetes (CFB), vanc increases firms and then proteo (with increasing dose)
•	Reeves/theriot/robinson—clindamycin/cef/cocktail/antbiotic administration alters the community structure
•	Croswell—prolonge impact of abx on intestinal microbial ecology salmonella
•	Rea- effect of  broad and narrow spectrum antimicrobials
•	Ferreira—intestinal microbiota plays a role in salmonella induced colitis
•	Theriot?—at least the metabolome paper—before and after with metabolites and cef, also does she look at anything taxonomically? yup

Also functional redundancy with a highly diverse community… not just about who possibly but also ab being diverse—emphasizes community as a whole

@@
